\documentclass[calc1-main.tex]{subfiles}

\begin{document}

\begin{example}
  How fast is the area of a rectangle changing if one side is 10cm long and is increasing at a rate of 2cm/s and the other side is 8cm long and is decreasing at a rate of 3cm/s?
\end{example}
\begin{solution}
  The area $A$, and the lengths of sides $x$ and $y$ are functions of time $t$. Also $A = x y$. We are given $\frac{dx}{dt} = 2$, $\frac{dy}{dt} = -3$ when $x=10$, $y=8$.

  Then
  \[
    \frac{dA}{dt} = \frac{dx}{dt} y + x \frac{dy}{dt}
  \]
  gives $\frac{dA}{dt} = -14$.
\end{solution}
In the previous problem, notice that the average changes
\[
  \begin{aligned}
    & \frac{A(1)-A(0)}{1-0} = 12\cdot 5 - 10 \cdot 8 = - 20, \\
    & \frac{A(.5)-A(0)}{.5-0} = 2(11\cdot 6.5 - 10 \cdot 8) = - 17 \\
    & \frac{A(.1)-A(0)}{.1-0} = 10(10.2\cdot 7.7 - 10 \cdot 8) = - 14.6
  \end{aligned}
\]
converge to the instantaneous rate we found.



This is possible because $A'(t) = -14 - 6t$. Hence $A$ changes in a non-constant fashion event though $x$ and $y$ changes constantly. What we are computing in this problem is $A'(0) = - 14$. And this result holds even if $\frac{dx}{dt}$ and $\frac{dy}{dt}$ are not constant. (maybe velocity of $x$ is not constant and it accelerates according to $x(t) = 10 + 2t + t^2$)

\begin{example}
  How fast is the surface area of a ball changing when the volume of the ball is $32 \pi/3$ cm$^3$ and is increasing at 2cm$^3/s$? (The surface are of the ball is $A = 4\pi r^2$ and the volume is $V = \frac{4}{3} \pi r^3$. Note that $V(r) = \int_0^r A(r) dr$)
\end{example}
\begin{solution}
  When $V=32 \pi/3$, $r=2$, $\frac{dV}{dt} = 2$
  \[
    \frac{dV}{dt} = 4 \pi r^2 \frac{dr}{dt}
  \]
  gives $\frac{dr}{dt} = \frac{1}{8 \pi}$.

  Now
  \[
    \frac{dA}{dt} = 8 \pi r \frac{dr}{dt} = 2
  \]
\end{solution}

\begin{example}
  A point is moving to the right along the first quadrant portion of the curve $x^2 y^3 = 72$. When the point has coordinates $(3,2)$, its horizontal velocity is 2 units/s. What is its vertical velocity?
\end{example}
\begin{solution}
  Taking $d/dt$ of both sides
  \[
    2x \frac{dx}{dt} y^3 + x^2 3 y^2 \frac{dy}{dt} = 0
  \]
  At $x=3$, $y=2$, $\frac{dx}{dt} = 2$,
  \[
    \frac{dy}{dt} = - \frac{8}{3}.
  \]

\end{solution}

\rule{\textwidth}{1pt}
\begin{multicols}{2}
\begin{exercise}
~\\
  \begin{enumerate}
    \item Find the rate of change of the area of a square whose side is 6 cm long, if the side length is increasing at 2 cm/min.
    \item Air is being pumped into a spherical balloon. The volume of the balloon is increasing at a rate of 20 cm$^3$/s when the radius is 30 cm. How fast is the radius increasing at that time? (The volume of a ball of radius r is $V=\frac{4}{3} \pi r^3$.)

    \item The area of a circle is decreasing at a rate of 2 cm$^2$/min. How fast is the radius of the circle changing when the area is 100 cm$^2$?

    Answer:
  \end{enumerate}
\end{exercise}
\end{multicols}
\rule{\textwidth}{1pt}
\end{document}
